---
title: "Sesame Final Project Report"
authors: "Julia Healey-Parera, Shelly Han, Angela Chen, Ayse Celebi, Amy Xu"
date: "November 27, 2025"
format: html
editor: visual
---

## Sta325 Final Project:

**1. Introduction**

Promoting early childhood learning should be a key priority in ensuring the quality of education and quality of life for individuals and families across the United States. Designed primarily for preschool-aged children, Sesame Street is one of the most widely studied early education television programs that focuses on foundational learning skills like early literacy and numeracy. This study analyzes trends in the Sesame Street viewing habits, socioeconomic background, and pre- and post-test performance of children in the U.S. aged 3-5 to explore whether exposure to educational television is associated with significant improvements in early education, and whether these effects differ across socioeconomic groups. Understanding these trends is not only important to families who hope to supplement their child's learning, but is also important to policymakers, media producers, and teachers in informing effective methods to design television series and curriculum or learning strategies to best support early education and potentially reduce early learning gaps caused by socioeconomic differences.

Our goals are to (i) evaluate which factors (e.g. age, socioeconomic status, and viewing frequency) are most influential in improving test scores, (ii) understand how socioeconomic status impacts preschool learning improvement and make predictions about these impacts, (iii) extend our methods to better understand the role of educational television in early childhood education using modern day educational examinations, and (iv) to apply our findings to inform future educational resource and curriculum design. We hypothesize that higher socioeconomic status and higher viewing frequency of Sesame Street are associated with greater increases in various tested areas.

Our objectives allow us to assess both the overall effectiveness of Sesame Street and how its impact may differ across populations. Our intended audiences benefit from both the statistical findings and the practical insights of our study. We hope that findings from this study are used in the future to better serve education in underserved communities, supporting efforts to improve education accessibility and equity, and overall, we hope this project contributes to understanding how accessible educational resources can complement formal preschool education.

**2. Methodology**

**2.1 Exploratory Data Analysis**

[Variables of Interest]{.underline}

Pre- and Post-tests in the following areas:

-   `body`: body parts (scores range from 0-32)

-   `let`: letters (scores range from 0-58)

-   `form`: forms (scores range from 0-20)

-   `numb`: numbers (scores range from 0-54)

-   `relat`: relational terms (scores range from 0-17)

-   `clasf`: classification skills (scores range from 0-24)

Since each test had a different range of scores, we standardized all pre- and post-test scores to be reported as a percentage of the total possible score, and we created a new *improvement* variable (e.g. `improve_body`, `improve_let`, etc.) for each of the tests to measure the percent improvement in each test area, which are our **Response Variables**.

**Covariates**:

-   `sex`: male = 1, female = 2

-   `age`: age in months

-   `peabody`: mental age score obtained from administration of the Peabody Picture Vocabulary test as a pretest measure of vocabulary maturity

-   `setting`: setting in which Sesame Street was viewed, 1=home 2=school

**Primary Predictor Variables**:

-   `viewenc`: treatment condition, 1=child encouraged to watch, 2=child not encouraged to watch

-   `viewcat`: frequency of viewing, 1=rarely watched the show, 2=once or twice a week, 3=three to five times a week, 4=watched the show on average more than 5 times a week

-   `site`: 1 =Three to five year old disadvantaged children from inner city areas in various parts of the country, 2 = Four year old advantaged suburban children, 3 = Advantaged rural children, 4 = Disadvantaged rural children, 5 = Disadvantaged Spanish speaking children.

```{r}
#| echo: False 
#| label: load-packages
#| message: False 
library(tidyverse)
library(ggplot2)
library(dplyr)
library(patchwork)
library(knitr)
library(rlang)
library(bamlss)
library(tidyr)
library(stringr)
```

```{r}
#| include: False
#| label: load-data
sesame <- read.csv("sesame.csv")
glimpse(sesame)
```

```{r}
#| include: False
#| label: data-cleaning
colnames(sesame)

colSums(is.na(sesame))
sum(!complete.cases(sesame))
nrow(sesame)

# standardizing test scores
max_scores <- c(
  body  = 32,
  let   = 58,
  form  = 20,
  numb  = 54,
  relat = 17, 
  clasf = 24
)

# convert scores to percent
sesame <- sesame |>
  mutate(
    prebody_p  = 100 * prebody  / max_scores["body"],
    postbody_p = 100 * postbody / max_scores["body"],

    prelet_p   = 100 * prelet   / max_scores["let"],
    postlet_p  = 100 * postlet  / max_scores["let"],

    preform_p  = 100 * preform  / max_scores["form"],
    postform_p = 100 * postform / max_scores["form"],

    prenumb_p  = 100 * prenumb  / max_scores["numb"],
    postnumb_p = 100 * postnumb / max_scores["numb"],

    prerelat_p  = 100 * prerelat  / max_scores["relat"],
    postrelat_p = 100 * postrelat / max_scores["relat"],
    
    preclasf_p  = 100 * preclasf  / max_scores["clasf"],
    postclasf_p = 100 * postclasf / max_scores["clasf"]
  )

# create variables for percent improvement for each test
sesame <- sesame |>
  mutate(
    improve_body  = postbody_p  - prebody_p,
    improve_let   = postlet_p   - prelet_p,
    improve_form  = postform_p  - preform_p,
    improve_numb  = postnumb_p  - prenumb_p,
    improve_relat = postrelat_p - prerelat_p,
    improve_clasf = postclasf_p - preclasf_p
  )
```

```{r}
#| echo: False
#| label: summary-stats-eda
# summary statistics
sesame |>
  select(
    improve_body,
    improve_let,
    improve_form,
    improve_numb,
    improve_relat,
    improve_clasf
  ) |>
  pivot_longer(cols = everything(),
               names_to = "Test",
               values_to = "Improvement") |>
  group_by(Test) |>
  summarize(
    Min = min(Improvement),
    Q1 = quantile(Improvement, 0.25),
    Median = median(Improvement),
    Mean = mean(Improvement),
    Q3 = quantile(Improvement, 0.75),
    Max = max(Improvement)
  ) |>
  kable(digits = 2)

# mean improvement by viewing category
sesame |>
  group_by(viewcat) |>
  summarize(
    body = mean(improve_body),
    letter = mean(improve_let),
    form = mean(improve_form),
    number = mean(improve_numb),
    relation = mean(improve_relat),
    classify = mean(improve_clasf)
  ) |>
  kable(
    digits = 2,
    caption = "Table 1: Mean Improvement by Viewing Category"
  )
```

The largest average percent improvements occurred in the tests for forms (19.06%), letters (18.63%), and numbers (16.95%), indicating the **strongest learning growth in forms, letters and numbers after Sesame Street exposure**. In contrast, there were only modest improvements in body knowledge (12.06%) and classification skills (14.60%) test scores and weak improvement in relational terms test scores (10.25%). All test had large negative minimums (as low as -64.81%) and high maximums (all at least 58.33%), indicating that there was greater variability in the improvements or lack therof in these tested areas for various students.

The mean improvement by viewing category table shows a **consistently strong positive relationship between viewing frequency and improvement across all test areas**, suggesting that Sesame Street exposure is associated with larger learning improvements. As viewing frequency increased from 1 to 4, **letter test scores showed the greatest increase** in percent improvement on average (from 4.28% to 27.75%).

```{r}
#| echo: False
#| label: univariate-eda
# boxplot distribution of all improvement scores
sesame |>
  select(
    improve_body, 
    improve_let, 
    improve_form, 
    improve_numb, 
    improve_relat, 
    improve_clasf
  ) |>
  pivot_longer(cols = everything()) |>
  ggplot(aes(x = name, y = value)) +
  geom_boxplot() +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  coord_flip() +
  labs(title = "Figure 1: Distributions of All Improvement Scores",
       x = "Skill Area",
       y = "Score Improvement")

# histogram of peabody distribution
hist_peabody <- ggplot(sesame, aes(peabody)) +
  geom_histogram(
    bins = 25,
    fill = "thistle3",
    color = "dimgrey",
    alpha = 0.8
  ) +
  labs(
    title = "Figure 2: Distribution of Peabody Vocabulary Scores",
    x = "Peabody Score",
    y = "Count"
  )

# histogram of improvement score for letters test
hist_let <- ggplot(sesame, aes(improve_let)) +
  geom_histogram(
    bins = 25, 
    fill = "lightskyblue3",
    color = "dimgrey"
  ) +
  labs(title = "Distribution of Letter Score Improvement",
       x = "Improvement (Post - Pre)",
       y = "Count")

# histogram of improvement score for numbers test
hist_numb <- ggplot(sesame, aes(improve_numb)) +
  geom_histogram(
    bins = 25, 
    fill = "lightskyblue3",
    color = "dimgrey"
  ) +
  labs(title = "Distribution of Numbers Score Improvement",
       x = "Improvement (Post - Pre)",
       y = "Count")

hist_peabody 
hist_let / hist_numb
```

Distributions of percent improvement across the 6 test areas showed **positive median improvement percentages for all 6 test areas**, with the largest median improvements *and* the greatest variability observed for the tests on **letters and numbers**. Most children demonstrated learning gains, but various outliers across almost all tests indicate variability in children's responses to Sesame Street exposure.

The distribution of Peabody vocabulary scores is **right-skewed**, indicating that most children are clustered in the lower score range and a smaller proportion of students achieve high Peabody vocabulary scores.

The distributions for letter and number test score improvements were both centered at values slightly above zero, indicating improvements on average, but both tests (especially for numbers) showed negative outliers, showing how some students experienced test score drops in numbers and letters after Sesame Street.

```{r}
#| echo: False
#| label: bivariate-eda-interactions

# interaction between viewcat and setting on improve_let
ggplot(sesame, aes(
  x = factor(viewcat),
  y = improve_let,
  fill = factor(setting)
)) +
  geom_boxplot() +
  scale_fill_manual(
    values = c("1" = "thistle3", "2" = "lightskyblue3"),
    labels = c("1" = "Home", "2" = "School")
  ) + 
  labs(
    title = "Figure 3: Interaction of Viewing Category and Educational Setting on Letter Improvement",
    x = "Viewing Category",
    y = "Letter Improvement",
    fill = "Educational Setting"
  )
```

Letter test score improvement increases consistently with viewing frequency for viewing at home (setting=1), but for viewing at a school setting (setting=2), letter score test percent improvement decreased at the highest viewing frequency, though not significantly. At **low-to-moderate viewing categories, viewing at school resulted in greater median letter test score improvements**, which at the highest viewing frequency, viewing at school resulted in a lower improvement than viewing at home; this could suggest that **too high of a viewing frequency in a school setting is actually detrimental to early learning**. Lastly, variability generally increases with higher viewing categories, showing that greater levels of Sesame Street exposure can lead to a larger variety of test outcomes for children.

Overall, the EDA shows that **Sesame Street exposure is positively associated with learning gains**, with the **largest improvements in forms, letters, and numbers**. **Viewing frequency displayed a positive relationship** with higher exposure linked to larger average improvements, especially for letters, though gains varied by setting at the highest viewing level. **All outcomes showed substantial variability and notable negative outliers**, indicating that while most children benefited, responses differed widely across individuals.

**2.2 Model Fitting**

```{r}
# Transformations 

sesame_edit <- sesame |> 
  mutate(
    # For each response, rescale (add 1 and divide by 2) to model with logit
    body_betatransf = (improve_body/100 + 1) / 2, 
    let_betatransf = (improve_let/100 + 1) / 2, 
    form_betatransf = (improve_form/100 + 1) / 2, 
    numb_betatransf = (improve_numb/100 + 1) / 2,
    relat_betatransf = (improve_relat/100 + 1) / 2, 
    clasf_betatransf = (improve_clasf/100 + 1) / 2,
    # Alt: for each response, rescale to approximate normal dist for tobit model
    # Indepedent variable transformations
    peabody_log = log(peabody),
    viewcat = factor(
      case_when(viewcat == 1 ~ 'rare', 
                viewcat == 2 ~ 'once_twice', 
                viewcat == 3 ~ 'three_five', 
                viewcat == 4 ~ 'over_five'), 
      levels = c('rare','once_twice','three_five', 'over_five')),
    site = factor(
      case_when(site == 1 ~ 'disadv_city', 
                site == 2 ~ 'adv_suburb', 
                site == 3 ~ 'adv_rural', 
                site == 4 ~ 'disadv_rural', 
                site == 5 ~ 'disadv_span'), 
      levels = c('disadv_span','disadv_city','disadv_rural','adv_rural','adv_suburb')), 
    sex = if_else(sex == 1, 'male', 'female'), 
    setting = factor(if_else(setting == 1, 'home', 'school'), levels = c('home', 'school')), 
    viewenc = if_else(viewenc == 1, 'encouraged', 'not_encouraged')
    )

write.csv(sesame_edit, 'sesame_edit.csv')
```

```{r}
# transformed response variables 
resp <- c("body_betatransf","let_betatransf","form_betatransf",
          "numb_betatransf","relat_betatransf","clasf_betatransf")

# starter formula for body (predictors) extraction to create all formulas
# start with all main effects and pairwise interaction terms, then perform selection afterwards
form_body <- as.formula(body_betatransf ~ la(site, type = 'single') + la(sex, type = 'single') + la(age, type = 'single') + la(viewcat, type = 'single') + la(setting, type = 'single') + la(viewenc, type = 'single') + la(peabody_log, type = 'single') + la(site : sex, type = 'single') + la(site : age, type = 'sing le') + la(site : viewcat, type = 'single') + la(site : viewenc, type = 'single') + la(site : peabody_log, type = 'single') + la(sex : age, type = 'single') + la(sex : viewcat, type = 'single') + la(sex : setting, type = 'single') + la(sex : viewenc, type = 'single') + la(sex : peabody_log, type = 'single') + la(age : viewcat, type = 'single') + la(age : setting, type = 'single') + la(age : viewenc, type = 'single') + + la(viewcat : setting, type = 'single') + la(viewcat : viewenc, type = 'single') + la(viewcat : peabody_log, type = 'single') + la(setting : viewenc, type = 'single') + la(setting : peabody_log, type = 'single') + la(viewenc : peabody_log, type = 'single'))

rhs_expr <- f_rhs(form_body)
formula_list <- list()

for (response_name in resp) {
  # take response_name string and make object for injection
  lhs_sym <- sym(response_name)
  
  new_formula <- new_formula(
    lhs = lhs_sym, 
    rhs = rhs_expr
  )
  formula_list[[response_name]] <- new_formula
}
```

```{r}
#| eval: false 

##### DO NOT RUN THIS CODE CHUNK -- takes roughly 20 minutes to train mods ######

model_list <- list()

for (curr_formula in formula_list) {
  print(curr_formula)
  currmod <- bamlss(
    curr_formula, 
    family = "beta", 
    data = sesame_edit, 
    optimizer = opt_lasso,
    criterion = "BIC", 
    nlambda = 100, 
    multiple = FALSE)
  
  currresponse <- f_lhs(curr_formula)
  model_list[[currresponse]] <- currmod
}

saveRDS(model_list[[1]], file = "models/initial/body_betamod.rds")
saveRDS(model_list[[2]], file = "models/initial/letters_betamod.rds")
saveRDS(model_list[[3]], file = "models/initial/form_betamod.rds")
saveRDS(model_list[[4]], file = "models/initial/numb_betamod.rds")
saveRDS(model_list[[5]], file = "models/initial/relat_betamod.rds")
saveRDS(model_list[[6]], file = "models/initial/clasf_betamod.rds")
```

```{r}
#| eval: false 

base_mod <- as.formula(body_betatransf ~ 
                         # main effects 
                         la(site, type = 'single') + 
                         la(sex, type = 'single') + 
                         la(age, type = 'single') + 
                         la(viewcat, type = 'single') + 
                         la(setting, type = 'single') + 
                         la(viewenc, type = 'single') + 
                         la(peabody_log, type = 'single') + 
                         # univerally significant interactions
                         la(site : peabody_log, type = 'single') + 
                         la(sex : setting, type = 'single') + 
                         la(sex : viewenc, type = 'single') + 
                         la(age : viewcat, type = 'single') + 
                         la(age : setting, type = 'single') + 
                         la(age : peabody_log, type = 'single') + 
                         la(setting : peabody_log, type = 'single') + 
                         la(viewenc : peabody_log, type = 'single')
                       )
  
body_mod <- update(base_mod, 
                   . ~ . + 
                     la(age : viewenc, type = 'single') +
                     la(viewcat : viewenc, type = 'single') +
                     la(sex : viewcat, type = 'single') +
                     la(viewcat : setting, type = 'single') +
                     la(setting : viewenc, type = 'single') +
                     la(sex : peabody_log, type = 'single')
                   )

let_mod <- update(base_mod, 
                  let_betatransf ~ . + 
                    la(site : age, type = 'single') +
                    la(sex : age, type = 'single') +
                    la(viewcat : peabody_log, type = 'single') +
                    la(sex : viewcat, type = 'single')  +
                    la(viewcat : setting, type = 'single') +
                    la(setting : viewenc, type = 'single') +
                    la(sex : peabody_log, type = 'single')
                  )

form_mod <- update(base_mod, 
                   form_betatransf ~ . + 
                     la(site : age, type = 'single') +
                     la(sex : age, type = 'single') +
                     la(viewcat : peabody_log, type = 'single') +
                     la(age : viewenc, type = 'single')  +
                     la(viewcat : viewenc, type = 'single') +
                     la(viewcat : setting, type = 'single') +
                     la(setting : viewenc, type = 'single') +
                     la(sex : peabody_log, type = 'single')
                   )

numb_mod <- update(base_mod, 
                   numb_betatransf ~ . + 
                     la(site : age, type = 'single') +
                     la(sex : age, type = 'single') +
                     la(viewcat : peabody_log, type = 'single') +
                     la(age : viewenc, type = 'single')  +
                     la(viewcat : viewenc, type = 'single') +
                     la(sex : viewcat, type = 'single')  +
                     la(setting : viewenc, type = 'single') +
                     la(sex : peabody_log, type = 'single')
                   )

relat_mod <- update(base_mod, 
                    relat_betatransf ~ . + 
                      la(site : age, type = 'single') + 
                      la(viewcat : peabody_log, type = 'single') +
                      la(viewcat : viewenc, type = 'single') +
                      la(sex : viewcat, type = 'single')  +
                      la(viewcat : setting, type = 'single') +
                      la(sex : peabody_log, type = 'single')
                    )

clasf_mod <- update(base_mod, 
                    clasf_betatransf ~ . + 
                      la(site : age, type = 'single') +
                      la(sex : age, type = 'single') +
                      la(viewcat : peabody_log, type = 'single') +
                      la(age : viewenc, type = 'single')  +
                      la(viewcat : viewenc, type = 'single') +
                      la(sex : viewcat, type = 'single')  +
                      la(setting : viewenc, type = 'single')
                    )

mods <- list(body_mod, let_mod, form_mod, numb_mod, relat_mod, clasf_mod)
```

```{r}
#| eval: false 

adj_models <- list()

for (mod in mods) {
  curr_adjmod <- bamlss(
    mod, 
    family = "beta", 
    data = sesame_edit, 
    optimizer = opt_lasso,
    criterion = "BIC", 
    nlambda = 100, 
    multiple = FALSE
    )
  
  adj_models[[f_lhs(mod)]] <- curr_adjmod
}

saveRDS(adj_models[[1]], file = "models/adjusted/body_adj_betamod.rds")
saveRDS(adj_models[[2]], file = "models/adjusted/letters_adj_betamod.rds")
saveRDS(adj_models[[3]], file = "models/adjusted/form_adj_betamod.rds")
saveRDS(adj_models[[4]], file = "models/adjusted/numb_adj_betamod.rds")
saveRDS(adj_models[[5]], file = "models/adjusted/relat_adj_betamod.rds")
saveRDS(adj_models[[6]], file = "models/adjusted/clasf_adj_betamod.rds")
```

```{r}
#| eval: false

### INITIAL MODEL TRAINING ERROR VALUES

rmse_body_initial <- mean((predict(model_list[[1]], model = "mu", type = "parameter") - sesame_edit$body_betatransf)^2)^(1/2)
rmse_letters_initial <- mean((predict(model_list[[2]], model = "mu", type = "parameter") - sesame_edit$let_betatransf)^2)^(1/2)
rmse_form_initial <- mean((predict(model_list[[3]], model = "mu", type = "parameter") - sesame_edit$form_betatransf)^2)^(1/2)
rmse_numb_initial <- mean((predict(model_list[[4]], model = "mu", type = "parameter") - sesame_edit$numb_betatransf)^2)^(1/2)
rmse_relat_initial <- mean((predict(model_list[[5]], model = "mu", type = "parameter") - sesame_edit$relat_betatransf)^2)^(1/2)
rmse_clasf_initial <- mean((predict(model_list[[6]], model = "mu", type = "parameter") - sesame_edit$clasf_betatransf)^2)^(1/2)

print(paste('Initial body model RMSE:', round(rmse_body_initial, 3)))
print(paste('Initial letters model RMSE:', round(rmse_letters_initial,3)))
print(paste('Initial form model RMSE:', round(rmse_form_initial,3)))
print(paste('Initial numbers model RMSE:', round(rmse_numb_initial,3)))
print(paste('Initial relational model RMSE:', round(rmse_relat_initial,3)))
print(paste('Initial classification model RMSE:', round(rmse_clasf_initial,3)))
```

```{r}
#| eval: false

### ADJUSTED MODEL TRAINING VALUES 

rmse_body_adj <- mean((predict(adj_models[[1]], model = "mu", type = "parameter") - sesame_edit$body_betatransf)^2)^(1/2)
rmse_letters_adj <- mean((predict(adj_models[[2]], model = "mu", type = "parameter") - sesame_edit$let_betatransf)^2)^(1/2)
rmse_form_adj <- mean((predict(adj_models[[3]], model = "mu", type = "parameter") - sesame_edit$form_betatransf)^2)^(1/2)
rmse_numb_adj <- mean((predict(adj_models[[4]], model = "mu", type = "parameter") - sesame_edit$numb_betatransf)^2)^(1/2)
rmse_relat_adj <- mean((predict(adj_models[[5]], model = "mu", type = "parameter") - sesame_edit$relat_betatransf)^2)^(1/2)
rmse_clasf_adj <- mean((predict(adj_models[[6]], model = "mu", type = "parameter") - sesame_edit$clasf_betatransf)^2)^(1/2)

print(paste('Adjusted body model RMSE:', round(rmse_body_adj, 3)))
print(paste('Adjusted letters model RMSE:', round(rmse_letters_adj,3)))
print(paste('Adjusted form model RMSE:', round(rmse_form_adj,3)))
print(paste('Adjusted numbers model RMSE:', round(rmse_numb_adj,3)))
print(paste('Adjusted relational model RMSE:', round(rmse_relat_adj,3)))
print(paste('Adjusted classification model RMSE:', round(rmse_clasf_adj,3)))
```

```{r}
#| eval: false

# Observation: all training errors are lower for adjusted models except for form and numbers, so rewrite adjusted models for initial form and number models and keep all others adjusted

saveRDS(adj_models[[1]], file = "models/final/body_fin_betamod.rds")
saveRDS(adj_models[[2]], file = "models/final/letters_fin_betamod.rds")
saveRDS(model_list[[3]], file = "models/final/form_fin_betamod.rds")
saveRDS(model_list[[4]], file = "models/final/numb_fin_betamod.rds")
saveRDS(adj_models[[5]], file = "models/final/relat_fin_betamod.rds")
saveRDS(adj_models[[6]], file = "models/final/clasf_fin_betamod.rds")
```

```{r}
### RUN THIS TO LOAD MODELS ###

body_final <- readRDS("models/final/body_fin_betamod.rds")
letters_final <- readRDS( "models/final/letters_fin_betamod.rds")
form_final <- readRDS("models/final/form_fin_betamod.rds")
numb_final <- readRDS("models/final/numb_fin_betamod.rds")
relat_final <- readRDS("models/final/relat_fin_betamod.rds")
clasf_final <- readRDS("models/final/clasf_fin_betamod.rds")
```

```{r}
# BIC values

print(paste('Body model BIC:', round(BIC(body_final),3)))
print(paste('Letters model BIC:', round(BIC(letters_final),3)))
print(paste('Form model BIC:', round(BIC(form_final),3)))
print(paste('Numbers model BIC:', round(BIC(numb_final),3)))
print(paste('Relational model BIC:', round(BIC(relat_final),3)))
print(paste('Classification model BIC:', round(BIC(clasf_final),3)))

rmse_body_final <- mean((predict(body_final, model = "mu", type = "parameter") - sesame_edit$body_betatransf)^2)^(1/2)
rmse_letters_final <- mean((predict(letters_final, model = "mu", type = "parameter") - sesame_edit$let_betatransf)^2)^(1/2)
rmse_form_final <- mean((predict(form_final, model = "mu", type = "parameter") - sesame_edit$form_betatransf)^2)^(1/2)
rmse_numb_final <- mean((predict(numb_final, model = "mu", type = "parameter") - sesame_edit$numb_betatransf)^2)^(1/2)
rmse_relat_final <- mean((predict(relat_final, model = "mu", type = "parameter") - sesame_edit$relat_betatransf)^2)^(1/2)
rmse_clasf_final <- mean((predict(clasf_final, model = "mu", type = "parameter") - sesame_edit$clasf_betatransf)^2)^(1/2)

print(paste('Final body model RMSE:', round(rmse_body_final, 3)))
print(paste('Final letters model RMSE:', round(rmse_letters_final,3)))
print(paste('Final form model RMSE:', round(rmse_form_final,3)))
print(paste('Final numbers model RMSE:', round(rmse_numb_final,3)))
print(paste('Final relational model RMSE:', round(rmse_relat_final,3)))
print(paste('Final classification model RMSE:', round(rmse_clasf_final,3)))
```

```{r}
# Graveyard                          
#la(site : viewcat, type = 'single')  # all
#la(site : viewenc, type = 'single') # all
#la(site : sex, type = 'single')  # all 
#la(site : age, type = 'single') # body
#la(sex : age, type = 'single') # body, relat
#la(viewcat : peabody_log, type = 'single') # body 
#la(age : viewenc, type = 'single')  # let, relat
#la(viewcat : viewenc, type = 'single') #let
#la(sex : viewcat, type = 'single')  # form
#la(viewcat : setting, type = 'single') # numb, clasf
#la(setting : viewenc, type = 'single') # relat
#la(sex : peabody_log, type = 'single') # clasf
```

**3. Results**

**3.1 Final Selected Models**

```{r}

form_initial <- readRDS("models/initial/form_betamod.rds")
numb_initial <- readRDS("models/initial/numb_betamod.rds")

model_objects <- list(
  body    = body_final,
  letters = letters_final,
  form    = form_initial,   
  numb    = numb_initial,   
  relat   = relat_final,
  clasf   = clasf_final
)

extract_la_terms_safe <- function(mod) {
  ftxt <- paste(deparse(formula(mod)), collapse = " ")
  matches <- str_match_all(ftxt,
      "la\\s*\\(\\s*([^,\\)]+)\\s*,\\s*type\\s*=\\s*['\"]?single['\"]?\\s*\\)")
  if (length(matches) == 0 || nrow(matches[[1]]) == 0) return(character(0))
  terms <- trimws(matches[[1]][,2])
  unique(terms)
}

terms_by_model <- lapply(model_objects, extract_la_terms_safe)
for (nm in names(terms_by_model)) {
  cat("\n---", nm, "(", length(terms_by_model[[nm]]), "terms) ---\n")
  if (length(terms_by_model[[nm]]) == 0) cat("(no la(...) terms detected)\n")
  else cat(paste0(terms_by_model[[nm]], collapse = "\n"), "\n")
}

all_terms <- sort(unique(unlist(terms_by_model)))
presence_tbl <- expand.grid(Predictor = all_terms, Model = names(model_objects),
                            stringsAsFactors = FALSE) |>
  rowwise() |>
  mutate(Kept = ifelse(Predictor %in% terms_by_model[[Model]], "✔", "✖")) |>
  ungroup() |>
  pivot_wider(names_from = Model, values_from = Kept)

cat("\n\n--- Predictor retention table ---\n")
print(kable(presence_tbl, caption = "Table 2: Predictors Kept Across Final Models"))

```

Table X summarizes the predictors kept in each of the six final beta regression models following the lasso-based variable-selection procedure. Across all outcomes, the main effects of age, site, viewing frequency (viewcat), viewing setting, viewing encouragement, and baseline vocabulary (Peabody score) were consistently retained, indicating that children's learning gains are influenced by a combination of developmental factors, socioeconomic context, and exposure to Sesame Street.

Several interaction terms also appeared in all or nearly all models, most notably interactions involving age, viewing frequency, viewing setting, and baseline vocabulary. The repeated selection of these interactions suggests that the effect of Sesame Street exposure is not uniform across children, but instead depends on how viewing behavior interacts with a child's age, learning environment, and prior vocabulary level. Some interactions, such as those involving sex or encouragement, appear in most but not all models. This indicates that their importance varies by outcome domain.

Overall, this retention pattern shows that children’s improvement is shaped not only by individual predictors but also by how those predictors interact, justifying the inclusion of interaction terms and supporting a nuanced interpretation of Sesame Street’s educational impact rather than a single average effect.

```{r}
# age & viewing frequency interaction effect
ggplot(sesame_edit, aes(x = age, y = improve_let, color = viewcat)) +
  geom_point(alpha = 0.25) +
  geom_smooth(se = FALSE) +
  scale_color_manual(values = c(
    "rare" = "thistle3",
    "once_twice" = "lightskyblue3",
    "three_five" = "darkseagreen3",
    "over_five" = "#D8A7B1"
  )) +
  labs(
    title = "Plot X: Age × Viewing Frequency on Letter Improvement",
    x = "Age (months)",
    y = "Letter Improvement (%)",
    color = "Viewing\nFrequency"
  ) +
  theme_minimal()

# age & viewing frequency interaction effect (Number Improvement)
ggplot(sesame_edit, aes(x = age, y = improve_numb, color = viewcat)) +
  geom_point(alpha = 0.25) +
  geom_smooth(se = FALSE) +
  scale_color_manual(values = c(
    "rare" = "thistle3",
    "once_twice" = "lightskyblue3",
    "three_five" = "darkseagreen3",
    "over_five" = "#D8A7B1"
  )) +
  labs(
    title = "Plot X: Age × Viewing Frequency on Number Improvement",
    x = "Age (months)",
    y = "Number Improvement (%)",
    color = "Viewing\nFrequency"
  ) +
  theme_minimal()


ggplot(sesame_edit, aes(x = age, y = improve_form, color = viewcat)) +
  geom_point(alpha = 0.25) +
  geom_smooth(se = FALSE) +
  scale_color_manual(values = c(
    "rare" = "thistle3",
    "once_twice" = "lightskyblue3",
    "three_five" = "darkseagreen3",
    "over_five" = "#D8A7B1"
  )) +
  labs(
    title = "Age × Viewing Frequency on Form Improvement",
    x = "Age (months)",
    y = "Form Improvement (%)",
    color = "Viewing\nFrequency"
  ) +
  theme_minimal()

ggplot(sesame_edit, aes(x = age, y = improve_relat, color = viewcat)) +
  geom_point(alpha = 0.25) +
  geom_smooth(se = FALSE) +
  scale_color_manual(values = c(
    "rare" = "thistle3",
    "once_twice" = "lightskyblue3",
    "three_five" = "darkseagreen3",
    "over_five" = "#D8A7B1"
  )) +
  labs(
    title = "Age × Viewing Frequency on Relations Improvement",
    x = "Age (months)",
    y = "Relations Improvement (%)",
    color = "Viewing\nFrequency"
  ) +
  theme_minimal()

ggplot(sesame_edit, aes(x = age, y = improve_clasf, color = viewcat)) +
  geom_point(alpha = 0.25) +
  geom_smooth(se = FALSE) +
  scale_color_manual(values = c(
    "rare" = "thistle3",
    "once_twice" = "lightskyblue3",
    "three_five" = "darkseagreen3",
    "over_five" = "#D8A7B1"
  )) +
  labs(
    title = "Age × Viewing Frequency on Classification Improvement",
    x = "Age (months)",
    y = "Classification Improvement (%)",
    color = "Viewing\nFrequency"
  ) +
  theme_minimal()


```

Since the interaction between `age` and `viewing frequency` was retained in all six final models, we visualize this relationship using letter score improvement as a representative outcome. Plot X shows that the association between age and learning gains varies meaningfully across viewing frequency levels.

For children who watch Sesame Street rarely or once-twice per week, letter improvement generally increases with age, indicating steady gains as children grow older. In contrast, for children who watch the program three to five times or more per week, improvements rise with age up to approximately 55 months, after which gains begin to level off or decline.

This pattern suggests that the relationship between very frequent exposure and learning gains differs at older ages, indicating that increased exposure does not translate into consistently larger improvements across the entire age range.

```{r}

# setting & viewing frequency interaction effect
ggplot(sesame_edit, aes(x = viewcat, y = improve_let, fill = setting)) +
  geom_boxplot(outlier.color = "dimgrey") +
  scale_fill_manual(values = c(
    "home"   = "lightskyblue3",
    "school" = "thistle3"
  )) +
  labs(
    title = "Plot J: Viewing Frequency × Setting on Letter Improvement",
    x = "Viewing Frequency",
    y = "Letter Improvement (%)",
    fill = "Viewing\nSetting"
  ) +
  theme_minimal()

ggplot(sesame_edit, aes(x = viewcat, y = improve_numb, fill = setting)) +
  geom_boxplot(outlier.color = "dimgrey") +
  scale_fill_manual(values = c(
    "home"   = "lightskyblue3",
    "school" = "thistle3"
  )) +
  labs(
    title = "Viewing Frequency × Setting on Number Improvement",
    x = "Viewing Frequency",
    y = "Number Improvement (%)",
    fill = "Viewing\nSetting"
  ) +
  theme_minimal()

ggplot(sesame_edit, aes(x = viewcat, y = improve_form, fill = setting)) +
  geom_boxplot(outlier.color = "dimgrey") +
  scale_fill_manual(values = c(
    "home"   = "lightskyblue3",
    "school" = "thistle3"
  )) +
  labs(
    title = "Viewing Frequency × Setting on Form Improvement",
    x = "Viewing Frequency",
    y = "Form Improvement (%)",
    fill = "Viewing\nSetting"
  ) +
  theme_minimal()

ggplot(sesame_edit, aes(x = viewcat, y = improve_relat, fill = setting)) +
  geom_boxplot(outlier.color = "dimgrey") +
  scale_fill_manual(values = c(
    "home"   = "lightskyblue3",
    "school" = "thistle3"
  )) +
  labs(
    title = "Viewing Frequency × Setting on Relations Improvement",
    x = "Viewing Frequency",
    y = "Relations Improvement (%)",
    fill = "Viewing\nSetting"
  ) +
  theme_minimal()

ggplot(sesame_edit, aes(x = viewcat, y = improve_clasf, fill = setting)) +
  geom_boxplot(outlier.color = "dimgrey") +
  scale_fill_manual(values = c(
    "home"   = "lightskyblue3",
    "school" = "thistle3"
  )) +
  labs(
    title = "Viewing Frequency × Setting on Classification Improvement",
    x = "Viewing Frequency",
    y = "Classification Improvement (%)",
    fill = "Viewing\nSetting"
  ) +
  theme_minimal()


```

The effect of viewing frequency on letter improvement differs by viewing setting. At low viewing frequencies, median improvements are similar for children watching at home and at school, indicating minimal differences between settings when exposure is limited.

At moderate viewing frequencies, differences between settings widen, with school-based viewing associated with higher median improvements. At the highest viewing frequency, however, this pattern reverses: improvements continue to increase for children watching at home, while gains for children watching at school decline and become more variable.

Overall, this interaction suggests that the association between viewing frequency and learning gains depends on the viewing environment, supporting the inclusion of the `viewing frequency × setting` interaction in the final models.

```{r}
# sex & viewing frequency interaction effect
ggplot(sesame_edit, aes(x = viewcat, y = improve_let, fill = sex)) +
  geom_boxplot(alpha = 0.75) +
  scale_fill_manual(values = c("female" = "thistle3", "male" = "lightskyblue3")) +
  labs(
    title = "Plot B: Sex × Viewing Frequency on Letter Improvement",
    y = "Letter Improvement (%)",
    x = "Viewing Frequency",
    fill = "Sex"
  ) +
  theme_minimal()

ggplot(sesame_edit, aes(x = viewcat, y = improve_numb, fill = sex)) +
  geom_boxplot(alpha = 0.75) +
  scale_fill_manual(values = c("female" = "thistle3", "male" = "lightskyblue3")) +
  labs(
    title = "Sex × Viewing Frequency on Number Improvement",
    y = "Number Improvement (%)",
    x = "Viewing Frequency",
    fill = "Sex"
  ) +
  theme_minimal()

ggplot(sesame_edit, aes(x = viewcat, y = improve_form, fill = sex)) +
  geom_boxplot(alpha = 0.75) +
  scale_fill_manual(values = c("female" = "thistle3", "male" = "lightskyblue3")) +
  labs(
    title = "Sex × Viewing Frequency on Form Improvement",
    y = "Form Improvement (%)",
    x = "Viewing Frequency",
    fill = "Sex"
  ) +
  theme_minimal()

ggplot(sesame_edit, aes(x = viewcat, y = improve_relat, fill = sex)) +
  geom_boxplot(alpha = 0.75) +
  scale_fill_manual(values = c("female" = "thistle3", "male" = "lightskyblue3")) +
  labs(
    title = "Sex × Viewing Frequency on Relations Improvement",
    y = "Relations Improvement (%)",
    x = "Viewing Frequency",
    fill = "Sex"
  ) +
  theme_minimal()

ggplot(sesame_edit, aes(x = viewcat, y = improve_clasf, fill = sex)) +
  geom_boxplot(alpha = 0.75) +
  scale_fill_manual(values = c("female" = "thistle3", "male" = "lightskyblue3")) +
  labs(
    title = "Sex × Viewing Frequency on Classification Improvement",
    y = "Classification Improvement (%)",
    x = "Viewing Frequency",
    fill = "Sex"
  ) +
  theme_minimal()

```

Plot B illustrates the interaction between sex and viewing frequency on letter improvement. For both females and males, median letter score improvement generally increases with viewing frequency. Among females, improvement rises from rare to moderate viewing levels and then levels off at higher viewing frequencies, indicating relatively stable changes beyond moderate exposure. In contrast, among males, improvement shows a more consistent upward trend across viewing categories, with higher viewing frequencies associated with larger median improvements.

Although these patterns suggest subtle differences in how viewing frequency relates to letter improvement by sex, the distributions substantially overlap across groups, indicating that sex-based differences are modest relative to the overall association between viewing frequency and letter improvement.

```{r}
get_clean_coefs <- function(model) {
  cf <- coef(model)[, 1]            
  cf <- cf[!grepl("alpha|tau21|Intercept", names(cf))]
  return(cf)
}

body_coef <- get_clean_coefs(body_final)
letters_coef <- get_clean_coefs(letters_final)
form_coef <- get_clean_coefs(form_final)
numb_coef <- get_clean_coefs(numb_final)
relat_coef <- get_clean_coefs(relat_final)
clasf_coef <- get_clean_coefs(clasf_final)


body_letters <- outer(body_coef, letters_coef, `-`)
body_form <- outer(body_coef, form_coef, `-`)
body_numb <- outer(body_coef, numb_coef, `-`)
body_relat <- outer(body_coef, relat_coef, `-`)
body_clasf <- outer(body_coef, clasf_coef, `-`)

letters_form <- outer(letters_coef, form_coef, `-`)
letters_numb <- outer(letters_coef, numb_coef, `-`)
letters_relat <- outer(letters_coef, relat_coef, `-`)
letters_clasf <- outer(letters_coef, clasf_coef, `-`)

form_numb <- outer(form_coef, numb_coef, `-`)
form_relat <- outer(form_coef, relat_coef, `-`)
form_clasf <- outer(form_coef, clasf_coef, `-`)

numb_relat <- outer(numb_coef,  relat_coef, `-`)
numb_clasf <- outer(numb_coef,  clasf_coef, `-`)

relat_clasf <- outer(relat_coef, clasf_coef, `-`)
```

```{r}
mat_list <- list(
  body_letters = body_letters,
  body_form    = body_form,
  body_numb    = body_numb,
  body_relat   = body_relat,
  body_clasf   = body_clasf,
  letters_form = letters_form,
  letters_numb = letters_numb,
  letters_relat= letters_relat,
  letters_clasf= letters_clasf,
  form_numb    = form_numb,
  form_relat   = form_relat,
  form_clasf   = form_clasf,
  numb_relat   = numb_relat,
  numb_clasf   = numb_clasf,
  relat_clasf  = relat_clasf
)

# Conver tot numeric from list
safe_to_numeric_matrix <- function(M, matrixname = NA_character_) {
  if (is.data.frame(M)) M <- as.matrix(M)
  if (is.atomic(M) && !is.matrix(M)) {
    M <- as.matrix(M)
  }
  if (is.numeric(M)) return(M)

  el_is_list <- any(vapply(M, is.list, logical(1)))
  if (!el_is_list) {
    # Not lists, attempt a vectorized numeric coercion (may produce NAs)
    coerced <- suppressWarnings(as.numeric(M))
    if (all(is.na(coerced) & !is.na(as.character(M)))) {
      stop(sprintf("Matrix '%s' cannot be coerced to numeric.", matrixname))
    }
    return(matrix(coerced, nrow = nrow(M), ncol = ncol(M)))
  }

  n <- length(M)
  coerced_vec <- numeric(n)
  for (i in seq_len(n)) {
    el <- M[i]
    if (is.list(el)) {
      val <- el[[1]]
      if (is.atomic(val) && length(val) == 1) {
        coerced_val <- suppressWarnings(as.numeric(val))
        coerced_vec[i] <- coerced_val
      } 
    } else {
      # not a list; coerce directly
      coerced_val <- suppressWarnings(as.numeric(el))
      coerced_vec[i] <- coerced_val
    }
  }

  matrix(coerced_vec, nrow = nrow(M), ncol = ncol(M))
}

top_k_from_mat_abs <- function(M_raw, k = 3, matrixname = NA_character_) {
  M <- safe_to_numeric_matrix(M_raw, matrixname = matrixname)

  signed_all <- as.numeric(M)
  abs_all <- abs(signed_all)
  valid <- !is.na(abs_all)
  
  lin_idx_all <- seq_along(M)
  ordered_lin <- lin_idx_all[valid][order(abs_all[valid], decreasing = TRUE)]
  k_use <- min(k, length(ordered_lin))
  idx_lin <- ordered_lin[seq_len(k_use)]
  idx_mat <- arrayInd(idx_lin, dim(M))

  rn <- rownames(M)
  cn <- colnames(M)
  row_names <- rn[idx_mat[,1]]
  col_names <- cn[idx_mat[,2]]

  signed_vals <- signed_all[idx_lin]
  abs_vals_sel <- abs(signed_vals)

  data.frame(
    matrix_name = rep(matrixname, k_use),
    row_index = idx_mat[,1],
    col_index = idx_mat[,2],
    row_name = row_names,
    column_name = col_names,
    absolutevalue_max = abs_vals_sel,
    signed_max = signed_vals
  )
}

top_k_table_for_list <- function(mat_list, k = 3) {
  nm <- names(mat_list)
  res_list <- lapply(nm, function(nm_i) {
      top_k_from_mat_abs(mat_list[[nm_i]], k = k, matrixname = nm_i) 
  })

  result <- do.call(rbind, res_list)
}


result_table <- top_k_table_for_list(mat_list, k = 3)

#### TABLE EXPLANATION: 
# For each of the above matrices, return the top 3 largest values (in other words, return the entries for which the coefficients for the two models are the most different). Each row provides the matrix for which the max came from. In matrix body_letters, body - letters was performed. Each entry is then the amount greater (or smaller, if negative) the coefficient for body is than letters. Can change k if you want to get more or less entries.
```

```{r}
# 1) Are all entries present and named?
names(mat_list)

# 2) What are the classes of each entry?
sapply(mat_list, class)

# 3) Are they matrices?
sapply(mat_list, is.matrix)

# 4) Dimensions (NA for non-matrix)
sapply(mat_list, function(x) if (is.matrix(x)) paste(dim(x), collapse = "x") else NA)
```

Because beta regression does not have a traditional R², we computed a correlation-based pseudo-R², defined as the squared correlation between observed improvements and predicted means, to summarize overall model fit.

```{r}
get_pseudo_r2 <- function(model, y) {
  mu_hat <- predict(model, model = "mu", type = "parameter")
  r2 <- cor(mu_hat, y, use = "complete.obs")^2
  return(r2)
}

# Compute pseudo-R² with simple names
r2_body    <- get_pseudo_r2(body_final,   sesame_edit$body_betatransf)
r2_letters <- get_pseudo_r2(letters_final,sesame_edit$let_betatransf)
r2_forms   <- get_pseudo_r2(form_initial, sesame_edit$form_betatransf)
r2_numbers <- get_pseudo_r2(numb_initial, sesame_edit$numb_betatransf)
r2_relat   <- get_pseudo_r2(relat_final,  sesame_edit$relat_betatransf)
r2_class   <- get_pseudo_r2(clasf_final,  sesame_edit$clasf_betatransf)

# Put them in a nice, normal table
pseudo_r2_table <- data.frame(
  Outcome = c("Body", "Letters", "Forms", "Numbers", "Relational", "Classification"),
  Pseudo_R2 = c(r2_body, r2_letters, r2_forms, r2_numbers, r2_relat, r2_class)
)

pseudo_r2_table

```

```{r}
mod <- if (exists("letters_adj_betamod")) {
  letters_adj_betamod
} else if (exists("letters_final")) {
  letters_final
} else {
  stop("No letters model object found in environment.")
}

# Plot BIC vs log(lambda) for the mu component
# (pathplot is provided by bamlss)
pathplot(mod, which = "criterion", model = "mu",
         main = "Figure 4: BIC vs. log(lambda) For The Letters model", xlab = "log(lambda)", ylab = "BIC")

# try to mark the chosen lambda with a dashed vertical line if it's stored in the model
chosen_lambda <- NULL
possible_names <- c("lambda", "lambda_mu", "lam", "opt.lambda")
for (nm in possible_names) {
  if (!is.null(mod[[nm]])) { chosen_lambda <- mod[[nm]]; break }
  if (!is.null(mod$opt) && !is.null(mod$opt[[nm]])) { chosen_lambda <- mod$opt[[nm]]; break }
}
# pathplot usually draws the chosen lambda already; this is a safe extra marker
if (!is.null(chosen_lambda) && length(chosen_lambda)>0) {
  abline(v = log(chosen_lambda[1]), lty = 2, lwd = 1)
  legend("topright", legend = "Selected λ (BIC)", lty = 2, bty = "n")
}
```

idk if this is right? BIC was used to select the LASSO tuning parameter λ over a grid of values. The BIC curve for the letters model shows that the optimal λ is relatively large, indicating that strong regularization produces the best balance between goodness-of-fit and model complexity, with many weaker effects shrunk toward zero.

```{r}
model_list <- list(
  body    = body_final,       # adjusted
  letters = letters_final,    # adjusted
  form    = form_initial,     # INITIAL model
  numb    = numb_initial,     # INITIAL model
  relat   = relat_final,      # adjusted
  clasf   = clasf_final       # adjusted
)

get_mu_coefs <- function(mod) {
  cf <- tryCatch(coef(mod, model = "mu"), error = function(e) coef(mod))
  
  # cf can be a matrix, vector, or list$mu — handle all:
  if (is.list(cf) && "mu" %in% names(cf)) {
    mat <- cf$mu
    vec <- as.numeric(mat[,1])
    names(vec) <- rownames(mat)
  } else if (is.matrix(cf)) {
    vec <- as.numeric(cf[,1])
    names(vec) <- rownames(cf)
  } else if (is.numeric(cf) && !is.null(names(cf))) {
    vec <- cf
  } else {
    stop("Could not extract coefficients for this model.")
  }
  
  # Drop nuisance parameters
  vec <- vec[!grepl("alpha|tau|Intercept", names(vec), ignore.case = TRUE)]
  
  return(vec)
}

# extract top 3 coefficients per model
top3_list <- map(names(model_list), function(m) {
  cf <- get_mu_coefs(model_list[[m]])
  
  # pick top 3 BY ABSOLUTE VALUE
  top_terms <- names(sort(abs(cf), decreasing = TRUE))[1:3]
  
  tibble(
    model = m,
    term = top_terms,
    estimate = cf[top_terms],
    abs_est = abs(cf[top_terms])
  )
})

top3_table <- bind_rows(top3_list)

kable(top3_table, digits = 4,
      caption = "Table 4: Top 3 Most Influential Predictors in Each Final Model")

# extract top 5 coefficients per model
top5_list <- map(names(model_list), function(m) {
  cf <- get_mu_coefs(model_list[[m]])
  
  # pick top 5 BY ABSOLUTE VALUE
  top_terms <- names(sort(abs(cf), decreasing = TRUE))[1:5]
  
  tibble(
    model = m,
    term = top_terms,
    estimate = cf[top_terms],
    abs_est = abs(cf[top_terms])
  )
})

top5_table <- bind_rows(top5_list)

kable(top5_table, digits = 4,
      caption = "Top 5 Most Influential Predictors in Each Final Model")


```

```{r}
#site x encouragement

my_cols3 <- c(
  "Main1"      = "#D8A7B1",
  "Main2"      = "lightskyblue3",
  "Interaction"= "thistle3"
)

summarize_site_enc_effects <- function(coef_vec) {
  nms <- names(coef_vec)

  site_terms <- nms[grepl("site", nms, ignore.case=TRUE) &
                    !grepl("enc|viewenc", nms, ignore.case=TRUE) &
                     (!grepl("[:~]", nms)) ]

  enc_terms  <- nms[grepl("encour|viewenc|enc", nms, ignore.case=TRUE) &
                    !grepl("site", nms, ignore.case=TRUE) &
                     (!grepl("[:~]", nms)) ]

  int_terms  <- nms[grepl("site", nms, ignore.case=TRUE) &
                    grepl("enc|viewenc", nms, ignore.case=TRUE)]

    get_signed <- function(terms) {
    if (length(terms)==0) return(0)
    coef_vec[terms][which.max(abs(coef_vec[terms]))]   # signed
  }

  tibble(
    effect = c("Main1","Main2","Interaction"),
    estimate = c(
      get_signed(site_terms),
      get_signed(enc_terms),
      get_signed(int_terms)
    ),
    magnitude = abs(estimate)
  )

}

df_siteenc3 <- imap_dfr(models_to_use, \(mod,name)
  summarize_site_enc_effects(get_mu_coefs(mod)) |> mutate(model=name))

df_siteenc3$effect <- factor(df_siteenc3$effect,
                             levels = c("Main1","Main2","Interaction"))


df_siteenc3$model <- factor(df_siteenc3$model,
                            levels=c("Body","Letters","Form","Numbers","Relat","Classif"))

p_siteenc3 <- ggplot(df_siteenc3,
                     aes(x=model, y=magnitude, fill=effect)) +
  geom_col(position=position_dodge(width=.7), width=.6, color="grey40") +
    geom_text(aes(label = sprintf("%.3f", estimate)),
            position = position_dodge(width = .7),
            vjust = -0.6, size = 1.8) +
  scale_fill_manual(
  values = my_cols3,
  breaks = c("Main1","Main2","Interaction"),
  labels = c("Site","Encouragement","Interaction"),
  name = "Effect"
) +
  labs(title="Figure 5: Strength of Site × Encouragement Effects Across Models",
       x="Model", y="Coefficient magnitude (logit-scale)") +
  theme_minimal(base_size=13) +
  theme(legend.position="top")

p_siteenc3

```

```{r}

data.frame(
  model = c('body', 'body', 'body', 'letters', 'letters', 'letters', 'form', 'form',)
)
```

```{r}
# sex x viewcat
summarize_sex_view_effects <- function(coef_vec) {
  nms <- names(coef_vec)

  sex_terms  <- nms[ (grepl("^sex|\\bsex\\b", nms, ignore.case=TRUE)) &
                     (!grepl("viewcat", nms, ignore.case=TRUE)) &
                     (!grepl("[:~]", nms)) ]

  view_terms <- nms[ (grepl("viewcat", nms, ignore.case=TRUE)) &
                     (!grepl("\\bsex\\b|^sex", nms, ignore.case=TRUE)) &
                     (!grepl("[:~]", nms)) ]

  int_terms  <- nms[ grepl("sex", nms, ignore.case=TRUE) &
                     grepl("viewcat", nms, ignore.case=TRUE) ]

    get_signed <- function(terms) {
    if (length(terms)==0) return(0)
    coef_vec[terms][which.max(abs(coef_vec[terms]))]   # signed
  }

  tibble(
    effect = c("Main1","Main2","Interaction"),
    estimate = c(
      get_signed(sex_terms),
      get_signed(view_terms),
      get_signed(int_terms)
    ),
    magnitude = abs(estimate)
  )

}


df_sexview3 <- imap_dfr(models_to_use, \(mod,name)
  summarize_sex_view_effects(get_mu_coefs(mod)) |> mutate(model=name))

df_sexview3$effect <- factor(df_sexview3$effect,
                             levels = c("Main1","Main2","Interaction"))

df_sexview3$model <- factor(df_sexview3$model,
                            levels=c("Body","Letters","Form","Numbers","Relat","Classif"))

p_sexview3 <- ggplot(df_sexview3,
                     aes(x=model, y=magnitude, fill=effect)) +
  geom_col(position=position_dodge(width=.7), width=.6, color="grey40") +
    geom_text(aes(label = sprintf("%.3f", estimate)),
            position = position_dodge(width = .7),
            vjust = -0.6, size = 1.8) +
  scale_fill_manual(values=my_cols3,
                    breaks = c("Main1","Main2","Interaction"),
                    labels=c("Sex","View Frequency","Interaction")) +
  labs(title="Figure 6: Strength of Sex × Viewcat Effects Across Models",
       x="Model", y="Coefficient magnitude (logit-scale)") +
  theme_minimal(base_size=13) +
  theme(legend.position="top")

p_sexview3

```

```{r}
# setting x vocab
summarize_setting_peabody_effects <- function(coef_vec) {
  nms <- names(coef_vec)

  # MAIN setting terms: contain "setting" but NOT "peabody" and NOT interaction markers (':' or '~')
  set_terms <- nms[
    grepl("setting", nms, ignore.case = TRUE) &
    !grepl("peabody", nms, ignore.case = TRUE) &
    !grepl("[:~]", nms)
  ]

  # PEABODY terms: contain "peabody" but NOT "setting" and NOT interaction markers
  peabody_terms <- nms[
    grepl("peabody", nms, ignore.case = TRUE) &
    !grepl("setting", nms, ignore.case = TRUE) &
    !grepl("[:~]", nms)
  ]

  # INTERACTIONS: contain both 'setting' and 'peabody'
  int_terms <- nms[
    grepl("setting", nms, ignore.case = TRUE) &
    grepl("peabody", nms, ignore.case = TRUE)
  ]

    get_signed <- function(terms) {
    if (length(terms) == 0) return(0)
    coef_vec[terms][which.max(abs(coef_vec[terms]))]   # signed
  }

  tibble(
    effect = c("Main1","Main2","Interaction"),
    estimate = c(
      get_signed(set_terms),
      get_signed(peabody_terms),
      get_signed(int_terms)
    ),
    magnitude = abs(estimate)
  )

}


df_setpeabody3 <- imap_dfr(models_to_use, \(mod,name)
  summarize_setting_peabody_effects(get_mu_coefs(mod)) |> mutate(model=name))

df_setpeabody3$effect <- factor(df_setpeabody3$effect,
                             levels = c("Main1","Main2","Interaction"))

df_setpeabody3$model <- factor(df_setpeabody3$model,
                               levels=c("Body","Letters","Form","Numbers","Relat","Classif"))

p_setpeabody3 <- ggplot(df_setpeabody3,
                       aes(x=model, y=magnitude, fill=effect)) +
  geom_col(position=position_dodge(width=.7), width=.6, color="grey40") +
    geom_text(aes(label = sprintf("%.3f", estimate)),
            position = position_dodge(width = .7),
            vjust = -0.6, size = 1.8) +
  scale_fill_manual(values=my_cols3,
                    breaks = c("Main1","Main2","Interaction"),
                    labels=c("Setting","Peabody","Interaction")) +
  labs(title="Figure 7: Strength of Setting × Vocabulary Effects Across Models",
       x="Model", y="Coefficient magnitude (logit-scale)") +
  theme_minimal(base_size=13) +
  theme(legend.position="top")

p_setpeabody3

```

```{r}

get_rmse <- function(model, y) {
  preds <- predict(model, model = "mu", type = "parameter")
  sqrt(mean((preds - y)^2, na.rm = TRUE))
}

rmse_body    <- get_rmse(body_final,   sesame_edit$body_betatransf)
rmse_letters <- get_rmse(letters_final,sesame_edit$let_betatransf)
rmse_forms   <- get_rmse(form_initial,  sesame_edit$form_betatransf)    
rmse_numbers <- get_rmse(numb_initial,  sesame_edit$numb_betatransf)    
rmse_relat   <- get_rmse(relat_final,  sesame_edit$relat_betatransf)
rmse_class   <- get_rmse(clasf_final,  sesame_edit$clasf_betatransf)

rmse_table <- data.frame(
  Model = c("Body",
            "Letters",
            "Forms",
            "Numbers",
            "Relational",
            "Classification"),
  RMSE = c(rmse_body, rmse_letters, rmse_forms, rmse_numbers, rmse_relat, rmse_class)
)

kable(rmse_table, digits = 3, 
      caption = "Table 3: Training RMSE for Final Models")


```

**4. Conclusion**

\pagebreak

**Sources**

**Appendix**
