---
title: "Sesame Final Project Report"
authors: "Julia Healey-Parera, Shelly Han, Angela Chen, Ayse Celebi, Amy Xu"
date: "November 27, 2025"
format: html
editor: visual
---

## Sta325 Final Project:

**1. Introduction**

Promoting early childhood learning should be a key priority in ensuring the quality of education and quality of life for individuals and families across the United States. Designed primarily for preschool-aged children, Sesame Street is one of the most widely studied early education television programs that focuses on foundational learning skills like early literacy and numeracy. This study analyzes trends in the Sesame Street viewing habits, socioeconomic background, and pre- and post-test performance of children in the U.S. aged 3-5 to explore whether exposure to educational television is associated with significant improvements in early education, and whether these effects differ across socioeconomic groups. Understanding these trends is not only important to families who hope to supplement their child's learning, but is also important to policymakers, media producers, and teachers in informing effective methods to design television series and curriculum or learning strategies to best support early education and potentially reduce early learning gaps caused by socioeconomic differences.

Our goals are to (i) evaluate which factors (e.g. age, socioeconomic status, and viewing frequency) are most influential in improving test scores, (ii) understand how socioeconomic status impacts preschool learning improvement and make predictions about these impacts, (iii) extend our methods to better understand the role of educational television in early childhood education using modern day educational examinations, and (iv) to apply our findings to inform future educational resource and curriculum design. We hypothesize that higher socioeconomic status and higher viewing frequency of Sesame Street are associated with greater increases in various tested areas.

Our objectives allow us to assess both the overall effectiveness of Sesame Street and how its impact may differ across populations. Our intended audiences benefit from both the statistical findings and the practical insights of our study. We hope that findings from this study are used in the future to better serve education in underserved communities, supporting efforts to improve education accessibility and equity, and overall, we hope this project contributes to understanding how accessible educational resources can complement formal preschool education.

**2. Methodology**

**2.1 Exploratory Data Analysis**

```{r}
#| echo: False 
#| label: load-packages
#| message: False 
library(tidyverse)
library(ggplot2)
library(dplyr)
library(patchwork)
library(knitr)
```

```{r}
#| include: False
#| label: load-data
sesame <- read.csv("sesame.csv")
glimpse(sesame)
```

```{r}
#| include: False
#| label: data-cleaning
colnames(sesame)

colSums(is.na(sesame))
sum(!complete.cases(sesame))
nrow(sesame)

# standardizing test scores
max_scores <- c(
  body  = 32,
  let   = 58,
  form  = 20,
  numb  = 54,
  relat = 17, 
  clasf = 24
)

# convert scores to percent
sesame <- sesame |>
  mutate(
    prebody_p  = 100 * prebody  / max_scores["body"],
    postbody_p = 100 * postbody / max_scores["body"],

    prelet_p   = 100 * prelet   / max_scores["let"],
    postlet_p  = 100 * postlet  / max_scores["let"],

    preform_p  = 100 * preform  / max_scores["form"],
    postform_p = 100 * postform / max_scores["form"],

    prenumb_p  = 100 * prenumb  / max_scores["numb"],
    postnumb_p = 100 * postnumb / max_scores["numb"],

    prerelat_p  = 100 * prerelat  / max_scores["relat"],
    postrelat_p = 100 * postrelat / max_scores["relat"],
    
    preclasf_p  = 100 * preclasf  / max_scores["clasf"],
    postclasf_p = 100 * postclasf / max_scores["clasf"]
  )

# create variables for percent improvement for each test
sesame <- sesame |>
  mutate(
    improve_body  = postbody_p  - prebody_p,
    improve_let   = postlet_p   - prelet_p,
    improve_form  = postform_p  - preform_p,
    improve_numb  = postnumb_p  - prenumb_p,
    improve_relat = postrelat_p - prerelat_p,
    improve_clasf = postclasf_p - preclasf_p
  )
```

```{r}
#| echo: False
#| label: summary-stats-eda
# summary statistics
sesame |>
  select(
    improve_body,
    improve_let,
    improve_form,
    improve_numb,
    improve_relat,
    improve_clasf
  ) |>
  pivot_longer(cols = everything(),
               names_to = "Test",
               values_to = "Improvement") |>
  group_by(Test) |>
  summarize(
    Min = min(Improvement),
    Q1 = quantile(Improvement, 0.25),
    Median = median(Improvement),
    Mean = mean(Improvement),
    Q3 = quantile(Improvement, 0.75),
    Max = max(Improvement)
  ) |>
  kable(digits = 2)

# mean improvement by viewing category
sesame |>
  group_by(viewcat) |>
  summarize(
    body = mean(improve_body),
    letter = mean(improve_let),
    form = mean(improve_form),
    number = mean(improve_numb),
    relation = mean(improve_relat),
    classify = mean(improve_clasf)
  ) |> 
  kable(digits = 2)
```

```{r}
#| echo: False
#| label: univariate-eda
# boxplot distribution of all improvement scores
sesame |>
  select(
    improve_body, 
    improve_let, 
    improve_form, 
    improve_numb, 
    improve_relat, 
    improve_clasf
  ) |>
  pivot_longer(cols = everything()) |>
  ggplot(aes(x = name, y = value)) +
  geom_boxplot() +
  coord_flip() +
  labs(title = "Distributions of All Improvement Scores",
       x = "Skill Area",
       y = "Score Improvement")

# histogram of peabody distribution
hist_peabody <- ggplot(sesame, aes(peabody)) +
  geom_histogram(
    bins = 25,
    fill = "thistle3",
    color = "dimgrey",
    alpha = 0.8
  ) +
  labs(
    title = "Distribution of Peabody Vocabulary Scores",
    x = "Peabody Score",
    y = "Count"
  )

# histogram of improvement score for letters test
hist_let <- ggplot(sesame, aes(improve_let)) +
  geom_histogram(
    bins = 25, 
    fill = "lightskyblue3",
    color = "dimgrey"
  ) +
  labs(title = "Distribution of Letter Score Improvement",
       x = "Improvement (Post - Pre)",
       y = "Count")

# histogram of improvement score for numbers test
hist_numb <- ggplot(sesame, aes(improve_numb)) +
  geom_histogram(
    bins = 25, 
    fill = "lightskyblue3",
    color = "dimgrey"
  ) +
  labs(title = "Distribution of Numbers Score Improvement",
       x = "Improvement (Post - Pre)",
       y = "Count")

hist_peabody 
hist_let / hist_numb
```

```{r}
#| echo: False
#| label: bivariate-eda-interactions

# interaction between viewcat and setting on improve_let
ggplot(sesame, aes(
  x = factor(viewcat),
  y = improve_let,
  fill = factor(setting)
)) +
  geom_boxplot() +
  labs(
    title = "Interaction of Treatment (Viewing Category) and SES on Letter Improvement",
    x = "Viewing Category (Treatment)",
    y = "Letter Improvement",
    fill = "Educational Setting (SES)"
  )
```

**2.2 Model Fitting**

**3. Results**

**3.1 Final Selected Models**

**3.3 Scientific Interpretation**

**4. Conclusion**

\pagebreak

**Sources**

**Appendix**
