---
title: "Sesame Final Project Report"
authors: "Julia Healey-Parera, Shelly Han, Angela Chen, Ayse Celebi, Amy Xu"
date: "November 27, 2025"
format: html
editor: visual
---

## Sta325 Final Project:

**1. Introduction**

Promoting early childhood learning should be a key priority in ensuring the quality of education and quality of life for individuals and families across the United States. Designed primarily for preschool-aged children, Sesame Street is one of the most widely studied early education television programs that focuses on foundational learning skills like early literacy and numeracy. This study analyzes trends in the Sesame Street viewing habits, socioeconomic background, and pre- and post-test performance of children in the U.S. aged 3-5 to explore whether exposure to educational television is associated with significant improvements in early education, and whether these effects differ across socioeconomic groups. Understanding these trends is not only important to families who hope to supplement their child's learning, but is also important to policymakers, media producers, and teachers in informing effective methods to design television series and curriculum or learning strategies to best support early education and potentially reduce early learning gaps caused by socioeconomic differences.

Our goals are to (i) evaluate which factors (e.g. age, socioeconomic status, and viewing frequency) are most influential in improving test scores, (ii) understand how socioeconomic status impacts preschool learning improvement and make predictions about these impacts, (iii) extend our methods to better understand the role of educational television in early childhood education using modern day educational examinations, and (iv) to apply our findings to inform future educational resource and curriculum design. We hypothesize that higher socioeconomic status and higher viewing frequency of Sesame Street are associated with greater increases in various tested areas.

Our objectives allow us to assess both the overall effectiveness of Sesame Street and how its impact may differ across populations. Our intended audiences benefit from both the statistical findings and the practical insights of our study. We hope that findings from this study are used in the future to better serve education in underserved communities, supporting efforts to improve education accessibility and equity, and overall, we hope this project contributes to understanding how accessible educational resources can complement formal preschool education.

**2. Methodology**

**2.1 Exploratory Data Analysis**

[Variables of Interest]{.underline}

Pre- and Post-tests in the following areas:

-   `body`: body parts (scores range from 0-32)

-   `let`: letters (scores range from 0-58)

-   `form`: forms (scores range from 0-20)

-   `numb`: numbers (scores range from 0-54)

-   `relat`: relational terms (scores range from 0-17)

-   `clasf`: classification skills (scores range from 0-24)

Since each test had a different range of scores, we standardized all pre- and post-test scores to be reported as a percentage of the total possible score, and we created a new *improvement* variable (e.g. `improve_body`, `improve_let`, etc.) for each of the tests to measure the percent improvement in each test area, which are our **Response Variables**.

**Covariates**:

-   `sex`: male = 1, female = 2

-   `age`: age in months

-   `peabody`: mental age score obtained from administration of the Peabody Picture Vocabulary test as a pretest measure of vocabulary maturity

-   `setting`: setting in which Sesame Street was viewed, 1=home 2=school

**Primary Predictor Variables**:

-   `viewenc`: treatment condition, 1=child encouraged to watch, 2=child not encouraged to watch

-   `viewcat`: frequency of viewing, 1=rarely watched the show, 2=once or twice a week, 3=three to five times a week, 4=watched the show on average more than 5 times a week

-   `site`: 1 =Three to five year old disadvantaged children from inner city areas in various parts of the country, 2 = Four year old advantaged suburban children, 3 = Advantaged rural children, 4 = Disadvantaged rural children, 5 = Disadvantaged Spanish speaking children.

```{r}
#| echo: False 
#| label: load-packages
#| message: False 
library(tidyverse)
library(ggplot2)
library(dplyr)
library(patchwork)
library(knitr)
library(rlang)
library(bamlss)
library(tidyr)
library(stringr)
```

```{r}
#| include: False
#| label: load-data
sesame <- read.csv("sesame.csv")
glimpse(sesame)
```

```{r}
#| include: False
#| label: data-cleaning
colnames(sesame)

colSums(is.na(sesame))
sum(!complete.cases(sesame))
nrow(sesame)

# standardizing test scores
max_scores <- c(
  body  = 32,
  let   = 58,
  form  = 20,
  numb  = 54,
  relat = 17, 
  clasf = 24
)

# convert scores to percent
sesame <- sesame |>
  mutate(
    prebody_p  = 100 * prebody  / max_scores["body"],
    postbody_p = 100 * postbody / max_scores["body"],

    prelet_p   = 100 * prelet   / max_scores["let"],
    postlet_p  = 100 * postlet  / max_scores["let"],

    preform_p  = 100 * preform  / max_scores["form"],
    postform_p = 100 * postform / max_scores["form"],

    prenumb_p  = 100 * prenumb  / max_scores["numb"],
    postnumb_p = 100 * postnumb / max_scores["numb"],

    prerelat_p  = 100 * prerelat  / max_scores["relat"],
    postrelat_p = 100 * postrelat / max_scores["relat"],
    
    preclasf_p  = 100 * preclasf  / max_scores["clasf"],
    postclasf_p = 100 * postclasf / max_scores["clasf"]
  )

# create variables for percent improvement for each test
sesame <- sesame |>
  mutate(
    improve_body  = postbody_p  - prebody_p,
    improve_let   = postlet_p   - prelet_p,
    improve_form  = postform_p  - preform_p,
    improve_numb  = postnumb_p  - prenumb_p,
    improve_relat = postrelat_p - prerelat_p,
    improve_clasf = postclasf_p - preclasf_p
  )
```

```{r}
#| echo: False
#| label: summary-stats-eda
# summary statistics
sesame |>
  select(
    improve_body,
    improve_let,
    improve_form,
    improve_numb,
    improve_relat,
    improve_clasf
  ) |>
  pivot_longer(cols = everything(),
               names_to = "Test",
               values_to = "Improvement") |>
  group_by(Test) |>
  summarize(
    Min = min(Improvement),
    Q1 = quantile(Improvement, 0.25),
    Median = median(Improvement),
    Mean = mean(Improvement),
    Q3 = quantile(Improvement, 0.75),
    Max = max(Improvement)
  ) |>
  kable(digits = 2)

# mean improvement by viewing category
sesame |>
  group_by(viewcat) |>
  summarize(
    body = mean(improve_body),
    letter = mean(improve_let),
    form = mean(improve_form),
    number = mean(improve_numb),
    relation = mean(improve_relat),
    classify = mean(improve_clasf)
  ) |> 
  kable(digits = 2)
```

The largest average percent improvements occurred in the tests for forms (19.06%), letters (18.63%), and numbers (16.95%), indicating the **strongest learning growth in forms, letters and numbers after Sesame Street exposure**. In contrast, there were only modest improvements in body knowledge (12.06%) and classification skills (14.60%) test scores and weak improvement in relational terms test scores (10.25%). All test had large negative minimums (as low as -64.81%) and high maximums (all at least 58.33%), indicating that there was greater variability in the improvements or lack therof in these tested areas for various students.

The mean improvement by viewing category table shows a **consistently strong positive relationship between viewing frequency and improvement across all test areas**, suggesting that Sesame Street exposure is associated with larger learning improvements. As viewing frequency increased from 1 to 4, **letter test scores showed the greatest increase** in percent improvement on average (from 4.28% to 27.75%).

```{r}
#| echo: False
#| label: univariate-eda
# boxplot distribution of all improvement scores
sesame |>
  select(
    improve_body, 
    improve_let, 
    improve_form, 
    improve_numb, 
    improve_relat, 
    improve_clasf
  ) |>
  pivot_longer(cols = everything()) |>
  ggplot(aes(x = name, y = value)) +
  geom_boxplot() +
  coord_flip() +
  labs(title = "Distributions of All Improvement Scores",
       x = "Skill Area",
       y = "Score Improvement")

# histogram of peabody distribution
hist_peabody <- ggplot(sesame, aes(peabody)) +
  geom_histogram(
    bins = 25,
    fill = "thistle3",
    color = "dimgrey",
    alpha = 0.8
  ) +
  labs(
    title = "Distribution of Peabody Vocabulary Scores",
    x = "Peabody Score",
    y = "Count"
  )

# histogram of improvement score for letters test
hist_let <- ggplot(sesame, aes(improve_let)) +
  geom_histogram(
    bins = 25, 
    fill = "lightskyblue3",
    color = "dimgrey"
  ) +
  labs(title = "Distribution of Letter Score Improvement",
       x = "Improvement (Post - Pre)",
       y = "Count")

# histogram of improvement score for numbers test
hist_numb <- ggplot(sesame, aes(improve_numb)) +
  geom_histogram(
    bins = 25, 
    fill = "lightskyblue3",
    color = "dimgrey"
  ) +
  labs(title = "Distribution of Numbers Score Improvement",
       x = "Improvement (Post - Pre)",
       y = "Count")

hist_peabody 
hist_let / hist_numb
```

Distributions of percent improvement across the 6 test areas showed **positive median improvement percentages for all 6 test areas**, with the largest median improvements *and* the greatest variability observed for the tests on **letters and numbers**. Most children demonstrated learning gains, but various outliers across almost all tests indicate variability in children's responses to Sesame Street exposure.

The distribution of Peabody vocabulary scores is **right-skewed**, indicating that most children are clustered in the lower score range and a smaller proportion of students achieve high Peabody vocabulary scores.

The distributions for letter and number test score improvements were both centered at values slightly above zero, indicating improvements on average, but both tests (especially for numbers) showed negative outliers, showing how some students experienced test score drops in numbers and letters after Sesame Street.

```{r}
#| echo: False
#| label: bivariate-eda-interactions

# interaction between viewcat and setting on improve_let
ggplot(sesame, aes(
  x = factor(viewcat),
  y = improve_let,
  fill = factor(setting)
)) +
  geom_boxplot() +
  labs(
    title = "Interaction of Viewing Category and Educational Setting on Letter Improvement",
    x = "Viewing Category",
    y = "Letter Improvement",
    fill = "Educational Setting"
  )
```

Letter test score improvement increases consistently with viewing frequency for viewing at home (setting=1), but for viewing at a school setting (setting=2), letter score test percent improvement decreased at the highest viewing frequency, though not significantly. At **low-to-moderate viewing categories, viewing at school resulted in greater median letter test score improvements**, which at the highest viewing frequency, viewing at school resulted in a lower improvement than viewing at home; this could suggest that **too high of a viewing frequency in a school setting is actually detrimental to early learning**. Lastly, variability generally increases with higher viewing categories, showing that greater levels of Sesame Street exposure can lead to a larger variety of test outcomes for children.

Overall, the EDA shows that **Sesame Street exposure is positively associated with learning gains**, with the **largest improvements in forms, letters, and numbers**. **Viewing frequency displayed a positive relationship** with higher exposure linked to larger average improvements, especially for letters, though gains varied by setting at the highest viewing level. **All outcomes showed substantial variability and notable negative outliers**, indicating that while most children benefited, responses differed widely across individuals.

**2.2 Model Fitting**

```{r}
# Transformations 

sesame_edit <- sesame |> 
  mutate(
    # For each response, rescale (add 1 and divide by 2) to model with logit
    body_betatransf = (improve_body/100 + 1) / 2, 
    let_betatransf = (improve_let/100 + 1) / 2, 
    form_betatransf = (improve_form/100 + 1) / 2, 
    numb_betatransf = (improve_numb/100 + 1) / 2,
    relat_betatransf = (improve_relat/100 + 1) / 2, 
    clasf_betatransf = (improve_clasf/100 + 1) / 2,
    # Alt: for each response, rescale to approximate normal dist for tobit model
    # Indepedent variable transformations
    peabody_log = log(peabody),
    viewcat = factor(
      case_when(viewcat == 1 ~ 'rare', 
                viewcat == 2 ~ 'once_twice', 
                viewcat == 3 ~ 'three_five', 
                viewcat == 4 ~ 'over_five'), 
      levels = c('rare','once_twice','three_five', 'over_five')),
    site = factor(
      case_when(site == 1 ~ 'disadv_city', 
                site == 2 ~ 'adv_suburb', 
                site == 3 ~ 'adv_rural', 
                site == 4 ~ 'disadv_rural', 
                site == 5 ~ 'disadv_span'), 
      levels = c('disadv_span','disadv_city','disadv_rural','adv_rural','adv_suburb')), 
    sex = if_else(sex == 1, 'male', 'female'), 
    setting = factor(if_else(setting == 1, 'home', 'school'), levels = c('home', 'school')), 
    viewenc = if_else(viewenc == 1, 'encouraged', 'not_encouraged')
    )

write.csv(sesame_edit, 'sesame_edit.csv')
```

```{r}
# transformed response variables 
resp <- c("body_betatransf","let_betatransf","form_betatransf",
          "numb_betatransf","relat_betatransf","clasf_betatransf")

# starter formula for body (predictors) extraction to create all formulas
# start with all main effects and pairwise interaction terms, then perform selection afterwards
form_body <- as.formula(body_betatransf ~ la(site, type = 'single') + la(sex, type = 'single') + la(age, type = 'single') + la(viewcat, type = 'single') + la(setting, type = 'single') + la(viewenc, type = 'single') + la(peabody_log, type = 'single') + la(site : sex, type = 'single') + la(site : age, type = 'sing le') + la(site : viewcat, type = 'single') + la(site : viewenc, type = 'single') + la(site : peabody_log, type = 'single') + la(sex : age, type = 'single') + la(sex : viewcat, type = 'single') + la(sex : setting, type = 'single') + la(sex : viewenc, type = 'single') + la(sex : peabody_log, type = 'single') + la(age : viewcat, type = 'single') + la(age : setting, type = 'single') + la(age : viewenc, type = 'single') + + la(viewcat : setting, type = 'single') + la(viewcat : viewenc, type = 'single') + la(viewcat : peabody_log, type = 'single') + la(setting : viewenc, type = 'single') + la(setting : peabody_log, type = 'single') + la(viewenc : peabody_log, type = 'single'))

rhs_expr <- f_rhs(form_body)
formula_list <- list()

for (response_name in resp) {
  # take response_name string and make object for injection
  lhs_sym <- sym(response_name)
  
  new_formula <- new_formula(
    lhs = lhs_sym, 
    rhs = rhs_expr
  )
  formula_list[[response_name]] <- new_formula
}
```

```{r}
#| eval: false 

##### DO NOT RUN THIS CODE CHUNK -- takes roughly 20 minutes to train mods ######

model_list <- list()

for (curr_formula in formula_list) {
  print(curr_formula)
  currmod <- bamlss(
    curr_formula, 
    family = "beta", 
    data = sesame_edit, 
    optimizer = opt_lasso,
    criterion = "BIC", 
    nlambda = 100, 
    multiple = FALSE)
  
  currresponse <- f_lhs(curr_formula)
  model_list[[currresponse]] <- currmod
}

saveRDS(model_list[[1]], file = "models/initial/body_betamod.rds")
saveRDS(model_list[[2]], file = "models/initial/letters_betamod.rds")
saveRDS(model_list[[3]], file = "models/initial/form_betamod.rds")
saveRDS(model_list[[4]], file = "models/initial/numb_betamod.rds")
saveRDS(model_list[[5]], file = "models/initial/relat_betamod.rds")
saveRDS(model_list[[6]], file = "models/initial/clasf_betamod.rds")
```

```{r}
#| eval: false 

base_mod <- as.formula(body_betatransf ~ 
                         # main effects 
                         la(site, type = 'single') + 
                         la(sex, type = 'single') + 
                         la(age, type = 'single') + 
                         la(viewcat, type = 'single') + 
                         la(setting, type = 'single') + 
                         la(viewenc, type = 'single') + 
                         la(peabody_log, type = 'single') + 
                         # univerally significant interactions
                         la(site : peabody_log, type = 'single') + 
                         la(sex : setting, type = 'single') + 
                         la(sex : viewenc, type = 'single') + 
                         la(age : viewcat, type = 'single') + 
                         la(age : setting, type = 'single') + 
                         la(age : peabody_log, type = 'single') + 
                         la(setting : peabody_log, type = 'single') + 
                         la(viewenc : peabody_log, type = 'single')
                       )
  
body_mod <- update(base_mod, 
                   . ~ . + 
                     la(age : viewenc, type = 'single') +
                     la(viewcat : viewenc, type = 'single') +
                     la(sex : viewcat, type = 'single') +
                     la(viewcat : setting, type = 'single') +
                     la(setting : viewenc, type = 'single') +
                     la(sex : peabody_log, type = 'single')
                   )

let_mod <- update(base_mod, 
                  let_betatransf ~ . + 
                    la(site : age, type = 'single') +
                    la(sex : age, type = 'single') +
                    la(viewcat : peabody_log, type = 'single') +
                    la(sex : viewcat, type = 'single')  +
                    la(viewcat : setting, type = 'single') +
                    la(setting : viewenc, type = 'single') +
                    la(sex : peabody_log, type = 'single')
                  )

form_mod <- update(base_mod, 
                   form_betatransf ~ . + 
                     la(site : age, type = 'single') +
                     la(sex : age, type = 'single') +
                     la(viewcat : peabody_log, type = 'single') +
                     la(age : viewenc, type = 'single')  +
                     la(viewcat : viewenc, type = 'single') +
                     la(viewcat : setting, type = 'single') +
                     la(setting : viewenc, type = 'single') +
                     la(sex : peabody_log, type = 'single')
                   )

numb_mod <- update(base_mod, 
                   numb_betatransf ~ . + 
                     la(site : age, type = 'single') +
                     la(sex : age, type = 'single') +
                     la(viewcat : peabody_log, type = 'single') +
                     la(age : viewenc, type = 'single')  +
                     la(viewcat : viewenc, type = 'single') +
                     la(sex : viewcat, type = 'single')  +
                     la(setting : viewenc, type = 'single') +
                     la(sex : peabody_log, type = 'single')
                   )

relat_mod <- update(base_mod, 
                    relat_betatransf ~ . + 
                      la(site : age, type = 'single') + 
                      la(viewcat : peabody_log, type = 'single') +
                      la(viewcat : viewenc, type = 'single') +
                      la(sex : viewcat, type = 'single')  +
                      la(viewcat : setting, type = 'single') +
                      la(sex : peabody_log, type = 'single')
                    )

clasf_mod <- update(base_mod, 
                    clasf_betatransf ~ . + 
                      la(site : age, type = 'single') +
                      la(sex : age, type = 'single') +
                      la(viewcat : peabody_log, type = 'single') +
                      la(age : viewenc, type = 'single')  +
                      la(viewcat : viewenc, type = 'single') +
                      la(sex : viewcat, type = 'single')  +
                      la(setting : viewenc, type = 'single')
                    )

mods <- list(body_mod, let_mod, form_mod, numb_mod, relat_mod, clasf_mod)
```

```{r}
#| eval: false 

adj_models <- list()

for (mod in mods) {
  curr_adjmod <- bamlss(
    mod, 
    family = "beta", 
    data = sesame_edit, 
    optimizer = opt_lasso,
    criterion = "BIC", 
    nlambda = 100, 
    multiple = FALSE
    )
  
  adj_models[[f_lhs(mod)]] <- curr_adjmod
}

saveRDS(adj_models[[1]], file = "models/adjusted/body_adj_betamod.rds")
saveRDS(adj_models[[2]], file = "models/adjusted/letters_adj_betamod.rds")
saveRDS(adj_models[[3]], file = "models/adjusted/form_adj_betamod.rds")
saveRDS(adj_models[[4]], file = "models/adjusted/numb_adj_betamod.rds")
saveRDS(adj_models[[5]], file = "models/adjusted/relat_adj_betamod.rds")
saveRDS(adj_models[[6]], file = "models/adjusted/clasf_adj_betamod.rds")
```

```{r}
#| eval: false

### INITIAL MODEL TRAINING ERROR VALUES

rmse_body_initial <- mean((predict(model_list[[1]], model = "mu", type = "parameter") - sesame_edit$body_betatransf)^2)^(1/2)
rmse_letters_initial <- mean((predict(model_list[[2]], model = "mu", type = "parameter") - sesame_edit$let_betatransf)^2)^(1/2)
rmse_form_initial <- mean((predict(model_list[[3]], model = "mu", type = "parameter") - sesame_edit$form_betatransf)^2)^(1/2)
rmse_numb_initial <- mean((predict(model_list[[4]], model = "mu", type = "parameter") - sesame_edit$numb_betatransf)^2)^(1/2)
rmse_relat_initial <- mean((predict(model_list[[5]], model = "mu", type = "parameter") - sesame_edit$relat_betatransf)^2)^(1/2)
rmse_clasf_initial <- mean((predict(model_list[[6]], model = "mu", type = "parameter") - sesame_edit$clasf_betatransf)^2)^(1/2)

print(paste('Initial body model RMSE:', round(rmse_body_initial, 3)))
print(paste('Initial letters model RMSE:', round(rmse_letters_initial,3)))
print(paste('Initial form model RMSE:', round(rmse_form_initial,3)))
print(paste('Initial numbers model RMSE:', round(rmse_numb_initial,3)))
print(paste('Initial relational model RMSE:', round(rmse_relat_initial,3)))
print(paste('Initial classification model RMSE:', round(rmse_clasf_initial,3)))
```

```{r}
#| eval: false

### ADJUSTED MODEL TRAINING VALUES 

rmse_body_adj <- mean((predict(adj_models[[1]], model = "mu", type = "parameter") - sesame_edit$body_betatransf)^2)^(1/2)
rmse_letters_adj <- mean((predict(adj_models[[2]], model = "mu", type = "parameter") - sesame_edit$let_betatransf)^2)^(1/2)
rmse_form_adj <- mean((predict(adj_models[[3]], model = "mu", type = "parameter") - sesame_edit$form_betatransf)^2)^(1/2)
rmse_numb_adj <- mean((predict(adj_models[[4]], model = "mu", type = "parameter") - sesame_edit$numb_betatransf)^2)^(1/2)
rmse_relat_adj <- mean((predict(adj_models[[5]], model = "mu", type = "parameter") - sesame_edit$relat_betatransf)^2)^(1/2)
rmse_clasf_adj <- mean((predict(adj_models[[6]], model = "mu", type = "parameter") - sesame_edit$clasf_betatransf)^2)^(1/2)

print(paste('Adjusted body model RMSE:', round(rmse_body_adj, 3)))
print(paste('Adjusted letters model RMSE:', round(rmse_letters_adj,3)))
print(paste('Adjusted form model RMSE:', round(rmse_form_adj,3)))
print(paste('Adjusted numbers model RMSE:', round(rmse_numb_adj,3)))
print(paste('Adjusted relational model RMSE:', round(rmse_relat_adj,3)))
print(paste('Adjusted classification model RMSE:', round(rmse_clasf_adj,3)))
```

```{r}
#| eval: false

# Observation: all training errors are lower for adjusted models except for form and numbers, so rewrite adjusted models for initial form and number models and keep all others adjusted

saveRDS(adj_models[[1]], file = "models/final/body_fin_betamod.rds")
saveRDS(adj_models[[2]], file = "models/final/letters_fin_betamod.rds")
saveRDS(model_list[[3]], file = "models/final/form_fin_betamod.rds")
saveRDS(model_list[[4]], file = "models/final/numb_fin_betamod.rds")
saveRDS(adj_models[[5]], file = "models/final/relat_fin_betamod.rds")
saveRDS(adj_models[[6]], file = "models/final/clasf_fin_betamod.rds")
```

```{r}
### RUN THIS TO LOAD MODELS ###

body_final <- readRDS("models/final/body_fin_betamod.rds")
letters_final <- readRDS( "models/final/letters_fin_betamod.rds")
form_final <- readRDS("models/final/form_fin_betamod.rds")
numb_final <- readRDS("models/final/numb_fin_betamod.rds")
relat_final <- readRDS("models/final/relat_fin_betamod.rds")
clasf_final <- readRDS("models/final/clasf_fin_betamod.rds")
```

```{r}
# BIC values

print(paste('Body model BIC:', round(BIC(body_final),3)))
print(paste('Letters model BIC:', round(BIC(letters_final),3)))
print(paste('Form model BIC:', round(BIC(form_final),3)))
print(paste('Numbers model BIC:', round(BIC(numb_final),3)))
print(paste('Relational model BIC:', round(BIC(relat_final),3)))
print(paste('Classification model BIC:', round(BIC(clasf_final),3)))

rmse_body_final <- mean((predict(body_final, model = "mu", type = "parameter") - sesame_edit$body_betatransf)^2)^(1/2)
rmse_letters_final <- mean((predict(letters_final, model = "mu", type = "parameter") - sesame_edit$let_betatransf)^2)^(1/2)
rmse_form_final <- mean((predict(form_final, model = "mu", type = "parameter") - sesame_edit$form_betatransf)^2)^(1/2)
rmse_numb_final <- mean((predict(numb_final, model = "mu", type = "parameter") - sesame_edit$numb_betatransf)^2)^(1/2)
rmse_relat_final <- mean((predict(relat_final, model = "mu", type = "parameter") - sesame_edit$relat_betatransf)^2)^(1/2)
rmse_clasf_final <- mean((predict(clasf_final, model = "mu", type = "parameter") - sesame_edit$clasf_betatransf)^2)^(1/2)

print(paste('Final body model RMSE:', round(rmse_body_final, 3)))
print(paste('Final letters model RMSE:', round(rmse_letters_final,3)))
print(paste('Final form model RMSE:', round(rmse_form_final,3)))
print(paste('Final numbers model RMSE:', round(rmse_numb_final,3)))
print(paste('Final relational model RMSE:', round(rmse_relat_final,3)))
print(paste('Final classification model RMSE:', round(rmse_clasf_final,3)))
```

```{r}
# Graveyard                          
#la(site : viewcat, type = 'single')  # all
#la(site : viewenc, type = 'single') # all
#la(site : sex, type = 'single')  # all 
#la(site : age, type = 'single') # body
#la(sex : age, type = 'single') # body, relat
#la(viewcat : peabody_log, type = 'single') # body 
#la(age : viewenc, type = 'single')  # let, relat
#la(viewcat : viewenc, type = 'single') #let
#la(sex : viewcat, type = 'single')  # form
#la(viewcat : setting, type = 'single') # numb, clasf
#la(setting : viewenc, type = 'single') # relat
#la(sex : peabody_log, type = 'single') # clasf
```

**3. Results**

**3.1 Final Selected Models**
```{r}


models <- c("body_final","letters_final","form_final","numb_final","relat_final","clasf_final")

extract_terms <- function(model_obj){
  # defensive: ensure object exists
  if (!exists(model_obj, envir = .GlobalEnv)) stop("Model not found: ", model_obj)
  f <- formula(get(model_obj, envir = .GlobalEnv))
  fs <- paste(deparse(f), collapse = " ")
  # Grab everything inside la(... , type = 'single' ) allowing single/double quotes and optional spaces
  m <- str_match_all(fs, "la\\s*\\(\\s*([^,\\)]+)\\s*,\\s*type\\s*=\\s*['\"]?single['\"]?")
  if (length(m[[1]]) == 0) return(character(0))
  terms <- str_trim(m[[1]][,2])
  unique(terms)
}

terms_by_model <- lapply(models, extract_terms)
names(terms_by_model) <- models

all_terms <- sort(unique(unlist(terms_by_model)))

presence_tbl <- expand.grid(
  Predictor = all_terms,
  Model = models,
  stringsAsFactors = FALSE
) %>%
  rowwise() %>%
  mutate(Kept = ifelse(Predictor %in% terms_by_model[[Model]], "✔", "✖")) %>%
  ungroup() %>%
  pivot_wider(names_from = Model, values_from = Kept)

kable(presence_tbl, caption = "Predictors Kept Across Final Models")

```




**3.3 Scientific Interpretation**

**4. Conclusion**

\pagebreak

**Sources**

**Appendix**
